---
title: "Algal Pond Community Composition"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE}
# libraries
library(ggplot2)

# INSERT PATH TO REPO ON COMPUTER HERE 
setwd("/Users/jonasgrove/Projects/pond_crashers/percent_composition_figure")

### HEALTHY PONDS
## algae52 ###### ###### ###### ###### ###### ######
algae_52_gtbtk_path <- "gtdbtk_algae52.tsv"
algae_52_checkm_path <- "profile_algae52.tsv"

# data
algae_52_gtbtk_data <- read.table(algae_52_gtbtk_path, sep = "\t")
algae_52_checkm_data <- read.table(algae_52_checkm_path, header = TRUE, sep = "\t")

# merge dataframes
algae52_combined <- merge(algae_52_gtbtk_data, algae_52_checkm_data, by.x = "V1", by.y = "Bin.Id")
colnames(algae52_combined) <- c("user_genome","domain", "phyllum", "class", "order", "family", "genus", "species","bin_size","mapped_reads","pecent_mapped","percent_binned","percent_community")

# sort dataframe
algae52_sorted <- algae52_combined[order(algae52_combined$percent_community, decreasing = TRUE),]

## algae53 ###### ###### ###### ###### ###### ######
algae_53_gtbtk_path <- "gtdbtk_algae53.tsv"
algae_53_checkm_path <- "profile_algae53.tsv"

# data
algae_53_gtbtk_data <- read.table(algae_53_gtbtk_path, sep = "\t")
algae_53_checkm_data <- read.table(algae_53_checkm_path, header = TRUE, sep = "\t")

# merge dataframes
algae53_combined <- merge(algae_53_gtbtk_data, algae_53_checkm_data, by.x = "V1", by.y = "Bin.Id")
colnames(algae53_combined) <- c("user_genome","domain", "phyllum", "class", "order", "family", "genus", "species","bin_size","mapped_reads","pecent_mapped","percent_binned","percent_community")

# sort dataframe
algae53_sorted <- algae53_combined[order(algae53_combined$percent_community, decreasing = TRUE),]



### SICK PONDS
## algae114 ###### ###### ###### ###### ###### ######
algae_114_gtbtk_path <- "gtdbtk_algae114.tsv"
algae_114_checkm_path <- "profile_algae114.tsv"

# data
algae_114_gtbtk_data <- read.table(algae_114_gtbtk_path, sep = "\t")
algae_114_checkm_data <- read.table(algae_114_checkm_path, header = TRUE, sep = "\t")

# merge dataframes
algae114_combined <- merge(algae_114_gtbtk_data, algae_114_checkm_data, by.x = "V1", by.y = "Bin.Id")
colnames(algae114_combined) <- c("user_genome","domain", "phyllum", "class", "order", "family", "genus", "species","bin_size","mapped_reads","pecent_mapped","percent_binned","percent_community")

# sort dataframe
algae114_sorted <- algae114_combined[order(algae114_combined$percent_community, decreasing = TRUE),]

## algae115 ###### ###### ###### ###### ###### ######
algae_115_gtbtk_path <- "gtdbtk_algae115.tsv"
algae_115_checkm_path <- "profile_algae115.tsv"

# data
algae_115_gtbtk_data <- read.table(algae_115_gtbtk_path, sep = "\t")
algae_115_checkm_data <- read.table(algae_115_checkm_path, header = TRUE, sep = "\t")

# merge dataframes
algae115_combined <- merge(algae_115_gtbtk_data, algae_115_checkm_data, by.x = "V1", by.y = "Bin.Id")
colnames(algae115_combined) <- c("user_genome","domain", "phyllum", "class", "order", "family", "genus", "species","bin_size","mapped_reads","pecent_mapped","percent_binned","percent_community")

# sort dataframe
algae115_sorted <- algae115_combined[order(algae115_combined$percent_community, decreasing = TRUE),]

#num_species <- 5

num_species <- 30
classification <- "phyllum"


species_comp <- function(num_species, classification){

  # algae 52 top two species and percentages
  algae52_tops <- algae52_sorted[1:num_species,c(classification, "percent_community")]
  
  # sum what percent of community was selected
  percent_selected <- 0
  for (i in seq(num_species)){
    percent_selected <- percent_selected + as.numeric(algae52_tops[i,"percent_community"])
  }
  # calculate and assign remaining percent community
  remaining <- 100 - percent_selected
  algae52_tops[num_species + 1, c(classification, "percent_community")] <- c("remaining", remaining) 
  algae52_tops$pond <- rep(c("Healthy1"), num_species + 1)
  
  # algae 53 top two species and percentages
  algae53_tops <- algae53_sorted[1:num_species,c(classification, "percent_community")]
  
  # sum what percent of community was selected
  percent_selected <- 0
  for (i in seq(num_species)){
    percent_selected <- percent_selected + as.numeric(algae53_tops[i,"percent_community"])
  }
  # calculate and assign remaining percent community
  remaining <- 100 - percent_selected
  algae53_tops[num_species + 1, c(classification, "percent_community")] <- c("remaining", remaining) 
  algae53_tops$pond <- rep(c("Healthy2"), num_species + 1)
  
  # algae 114 top two species and percentages
  algae114_tops <- algae114_sorted[1:num_species,c(classification, "percent_community")]
  
  # sum what percent of community was selected
  percent_selected <- 0
  for (i in seq(num_species)){
    percent_selected <- percent_selected + as.numeric(algae114_tops[i,"percent_community"])
  }
  # calculate and assign remaining percent community
  remaining <- 100 - percent_selected
  algae114_tops[num_species + 1, c(classification, "percent_community")] <- c("remaining", remaining) 
  algae114_tops$pond <- rep(c("Sick1"), num_species + 1)
  
  # algae 115 top two species and percentages
  algae115_tops <- algae115_sorted[1:num_species,c(classification, "percent_community")]
  
  # sum what percent of community was selected
  percent_selected <- 0
  for (i in seq(num_species)){
    percent_selected <- percent_selected + as.numeric(algae114_tops[i,"percent_community"])
  }
  # calculate and assign remaining percent community
  remaining <- 100 - percent_selected
  algae115_tops[num_species + 1, c(classification, "percent_community")] <- c("remaining", remaining) 
  algae115_tops$pond <- rep(c("Sick2"), num_species + 1)
  
  # combine sick and healthy ponds
  all_ponds <- data.frame()
  
  # add algae52 
  start <- 1
  end <- num_species + 1
  all_ponds[start:end, 1:3] <- algae52_tops
  
  # add algae53
  start <- end + 1
  end <- start + num_species 
  all_ponds[start:end, 1:3] <- algae53_tops
  
  # add algae114
  start <- end + 1
  end <- start + num_species 
  all_ponds[start:end,1:3] <- algae114_tops
  
  # algae 115
  start <- end + 1
  end <- start + num_species 
  all_ponds[start:end,1:3] <- algae115_tops
  
  # make scale 0-1
  all_ponds$percent_community <- as.numeric(all_ponds$percent_community)/100
    
  # Stacked + percent
  ggplot(all_ponds, aes_string(fill=classification, y="percent_community", x="pond")) + 
    geom_bar(position="fill", stat="identity") + 
    labs(title = "N. oceanica Pond Composition", y = "Percent Community", x = "Pond Sample") +
    theme(title = element_text(size = 25), 
          axis.title.x = element_text(size = 20), 
          axis.title.y = element_text(size = 20),
          axis.text.x = element_text(size = 15),
          axis.text.y = element_text(size = 15),
          legend.text = element_text(size = 15))

}
```





```{r echo=FALSE}
inputPanel(
  selectInput("classification", label = " Classification:",
              choices = c("domain", "phyllum", "class", "order", "family", "genus", "species"),
              selected = "phyllum"),
  selectInput("n_species", label = "Number of Bins:",
              choices = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30), 
              selected = 3))

renderPlot({
  species_comp(as.numeric(input$n_species), input$classification)
})
```

##### The above figure shows the percent community composition of each Nannochloropsis oceanica pond. Classification and quantification were determined by gtdbtk and checkm, respectively. If a classification option is chosen where some bins did not have sufficient resolution to meet the classification, those bins are put into the category **x\_\_**, where **x** is the chosen classification. The category **remaining** represents the percentage of pond composition that was not annotated in the analysis pipeline, or not included in the chosen number of bins. 


